---
title: "NHANES: What affects our sleep?"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
pacman::p_load(
    rio,
    here,
    tidyverse,
    flexdashboard,
    shiny,
    plotly
)

# import data
```

# Sidebar {.sidebar}
On average, humans spend about a 1/3 of each day sleeping. Good sleep is not only vital to our day-to-day functioning but also allows us to lead healthy lives. In the US, 50-70 million people suffer from insomnia. This dashboard uses the 2018 NHANES data to explore the different variables that can affect our sleep. 

Top 10 variables that affect sleep were identified using machine learning. Here, users can explore the data and run different machine learning models to learn about these variables in depth.

App authors: Briha Ansari, MD., Feng-Chiao Lee, Tim Lee, MD.

# Exploratory Data Analysis
Column  {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}

```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}

```

### Chart C

```{r}

```

# Logistic Regression

# Random Forest

# XGBoost
