---
title: "NHANES: What affects our sleep?"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    theme: "sandstone" # or"yeti"
    source_code: embed
runtime: shiny

---

```{r global, include=FALSE}
pacman::p_load(
    dplyr,
    tidyverse,
    flexdashboard,
    shiny,
    shinydashboard,
    plotly,
    summarytools,
    prettycode,
    titanic,
    lmerTest,
    ggridges,
    DT,
    ggeffects,
    splines,
    ggiraphExtra,
    ggplot2,
    ResourceSelection,
    pscl,
    DALEX,
    DALEXtra,
    tidymodels,
    pROC
)
# import data
dat <- load(file = "ready.rda")
final_dat <- ready
```

# Exploratory Data Analysis
Inputs {.sidebar}
-----------------------------------------------------------------------


**Exploratory Data Analysis**\
Change the X and Y to explore the relationships between various variables. "Color by group" segments the chart further by the selected group but will not work when 2 non-numeric variables are chosen.


```{r}
# separate factors and numeric variables
var_factor <- final_dat %>% 
  select_if(is.factor)
var_factor <- colnames(var_factor)
var_numeric <- final_dat %>% 
  select_if(is.numeric)
var_numeric <- colnames(var_numeric[-1])
# shiny app input numeric
selectInput("x", label = "X", names(final_dat))
selectInput("y", label = "Y", c("none", names(final_dat)), names(final_dat)[[2]])
selectInput("color", "Color by group", c("none", var_factor))
```


App authors: Briha Ansari, MD., Feng-Chiao Lee, Tim Lee, MD.

Row
-----------------------------------------------------------------------
### Total N
```{r}
total <- nrow(final_dat)
flexdashboard::valueBox(total, icon = "fa-users")
```

### Average sleep hours
```{r}
hours <- 7.6
flexdashboard::valueBox(hours, icon = "fa-bed")
```

### % Trouble sleeping
```{r}
flexdashboard::valueBox(scales::percent(.285), icon = "fa-exclamation-circle",
        color = "warning")
```

Row
-----------------------------------------------------------------------
### Plot

```{r}
# get plot type
# 2 = both numeric
# 1 = one numeric, one non-numeric
# 0 = both non-numeric
# -1 = only one variable
plot_type <- reactive({
  if(input$y != "none")
    is.numeric(final_dat[[input$x]]) + is.numeric(final_dat[[input$y]])
  else
    -1
})
renderPlot({
  if (plot_type() ==2) {
    # 2 num = scatterplot
    p <- ggplot(final_dat, aes_string(input$x, input$y)) +
      geom_point(alpha = 0.5) +
      geom_smooth()
    # color change
    if (input$color != "none")
      p <- p + aes_string(color = input$color)
  } else if (plot_type() == 1) {
    # 1 num, 1 non-num = boxplot
    p <- ggplot(final_dat, aes_string(input$x, input$y)) +
      geom_boxplot()
    
    # fill change
    if (input$color != "none")
      p <- p + aes_string(fill = input$color)
  } 
  else if (plot_type() == 0 ) {
    # 2 non-num = heatmap
    temp_dat <- reactive(final_dat[, c(input$x, input$y)] %>% 
                           group_by(across()) %>% 
                           summarize(proportion = n()/length(ready$sleep_problems))
                         )
    p <- ggplot(temp_dat(),
                mapping = aes_string(x = input$x, y = input$y, fill = "proportion")) +
      geom_tile()
  } else {
    # 1 var only = univariate plot
    p <- ggplot(final_dat, aes_string(x = input$x))
    
    if(is.numeric(final_dat[[input$x]]))
      p <- p + geom_histogram()
    else
      p <- p + geom_bar()
    
    # fill change
    if(input$color != "none")
      p <- p + aes_string(fill = input$color)
  }
  
  # add title
  if(plot_type() >= 0) {
    p <- p + labs(title = paste(input$y, "vs.", input$x))
  } else {
    p <- p + labs(title = paste("Distribution of", input$x))
  }
  
  # add styling
  p <- p +
    theme_bw() +
    theme(plot.title = element_text(size = rel(1.8), face = "bold", hjust = 0.5),
          axis.title = element_text(size = rel(1.2)))
  print(p)
})
```


# Logistic Regression

Inputs {.sidebar}
-----------------------------------------------------------------------
Answer the following questions to find out your likelihood of having sleep problems:

```{r}
# age
selectInput("age.cat2", label = "age.cat2: Your Age Range", 
    choices = list("19-35" = "19-35", "36-50" = "36-50","51-65"="51-65","above65"="above65"), 
    selected = "19-35")

# bmi
numericInput("bmi", label = "bmi: Body Mass Index (kg/m^2)", value = 20,  min = 12, max = 90, step = 0.1)  
```

**Mental Health**
```{r}
# mental health
# depression
selectInput("depression", label = "depression: How often do you feel depressed?", 
    choices = list("daily" = "Daily", "weekly" = "weekly", "monthly" = "monthly",'few_times_year'='few_times_year','never'='never'), 
    selected = "never")

# anxiety
selectInput("anxious", label = "anxious:  How often do you feel worried or anxious?", 
    choices = list("daily" = "daily", "monthly" = "monthly","few/yr"="few/yr","never"="never"), 
    selected = "never")
```

**Physical Health**
```{r}
# physical strength
selectInput("difficult_move_objects", label = "difficult_move_objects: Difficulty moving large objects?", 
    choices = list("no" = "no", "some" = "some", "much" = "much",'unable'='unable','dont_do_this_activity'='dont_do_this_activity'), 
    selected = "no")

# snorting
selectInput("snort", label = "snort:  How often do you snort or stop breathing?", 
    choices = list("never" = "never", "1-2/wk" = "1-2/wk","3-4/wk"="3-4/wk","5/wk"="5/wk"), 
    selected = "never")
```

**Lab values**
```{r}
# lab values
# crp
sliderInput("crp", label = "crp: HS C-Reactive Protein (mg/L)?",
            min = 0, 
            max = round(max(ready$crp),2), 
            value =round(0,2),step =0.01)

# iron
sliderInput("iron_um_l", label = "iron_um_l - Iron frozen, Serum (umol/L)",
            min =0, 
            max = round(max(ready$iron_um_l),2), 
            value =round(20,2),step =0.5)

```

**Plot**
```{r}
selectInput("variable", label = "Choose a Variable to see how it affects sleep problems", names(ready))
```


```{r reactive expression}
logitMod <- glm(formula = sleep_problems ~ depression  
                + bmi 
                + difficult_move_objects 
                + snort+crp+anxious
                + iron_um_l
                + age.cat2, data = ready,family = binomial(link = "logit"))
```


Row {data-height=450}
-----------------------------------------------------------------------
### Variable Effect on Sleeping Problem
```{r}
flexdashboard::renderValueBox({
  testdata <- data.frame(depression =factor(input$depression),
                       snort=factor(input$snort),
                       difficult_move_objects = factor(input$difficult_move_objects),
                       snort = factor(input$snort),
                       crp = input$crp,
                       iron_um_l = input$iron_um_l,
                       anxious = factor(input$anxious),
                       bmi=input$bmi,
                       age.cat2 = factor(input$age))
testdata$prob<- predict(logitMod, newdata=testdata, type="response")
flexdashboard::valueBox(scales::percent(round(testdata$prob, 2)),
         icon = "fa-bed")
})
```

### break down profile
```{r histogram}
renderPlot({
  
ggpredict(logitMod, terms = input$variable ,fill=input$variable)%>%plot()
  })
```


Row {data-height=450}
--------------
### Vip
```{r}
renderPlotly({
  library(vip)
  vip(logitMod, mapping = aes_string(fill = "Variable"))
  
})
```

### Performance
```{r}
library(pROC)
test_prob = predict(logitMod, newdata = final_dat, type = "response")
test_roc = roc(final_dat$sleep_problems ~ test_prob, plot = TRUE, print.auc = TRUE)
```




# Regularized Logit

-----------------------------------------------------------------------


```{r load regularized regression model & data}
log_reg <- readRDS("final_lr_wf.RDS")
#load("ready.rda")
```


```{r explainable, include=FALSE}
log_reg_trained <- log_reg %>%
  fit(data = final_dat)
log_reg_explainer <- explain_tidymodels(
                        model = log_reg_trained,
                        data = final_dat %>%
                          select(-sleep_problems),
                        y = ready$sleep_problems,
                        label = "Regularized Logistic Regression")
```
Inputs {.sidebar}
-----------------------------------------------------------------------


Answer the following questions to find out your likelihood of having sleep problems:

```{r}
# age
selectInput("age.cat2", label = "age.cat2: Your Age Range", 
    choices = list("19-35" = "19-35", "36-50" = "36-50","51-65"="51-65","above65"="above65"), 
    selected = "19-35")

# bmi
numericInput("bmi", label = "bmi: Body Mass Index (kg/m^2)", value = 20,  min = 12, max = 90, step = 0.1)  
```

**Mental Health**
```{r}
# mental health
# depression
selectInput("depression", label = "depression: How often do you feel depressed?", 
    choices = list("daily" = "Daily", "weekly" = "weekly", "monthly" = "monthly",'few_times_year'='few_times_year','never'='never'), 
    selected = "never")

# anxiety
selectInput("anxious", label = "anxious:  How often do you feel worried or anxious?", 
    choices = list("daily" = "daily", "monthly" = "monthly","few/yr"="few/yr","never"="never"), 
    selected = "never")
```

**Physical Health**
```{r}
# physical strength
selectInput("difficult_move_objects", label = "difficult_move_objects: Difficulty moving large objects?", 
    choices = list("no" = "no", "some" = "some", "much" = "much",'unable'='unable','dont_do_this_activity'='dont_do_this_activity'), 
    selected = "no")

# snorting
selectInput("snort", label = "snort:  How often do you snort or stop breathing?", 
    choices = list("never" = "never", "1-2/wk" = "1-2/wk","3-4/wk"="3-4/wk","5/wk"="5/wk"), 
    selected = "never")
```

**Lab values**
```{r}
# lab values
# crp
sliderInput("crp", label = "crp: HS C-Reactive Protein (mg/L)?",
            min = 0, 
            max = round(max(ready$crp),2), 
            value =round(0,2),step =0.01)

# iron
sliderInput("iron_um_l", label = "iron_um_l - Iron frozen, Serum (umol/L)",
            min =0, 
            max = round(max(ready$iron_um_l),2), 
            value =round(20,2),step =0.5)

```

```{r}
log_reg_obs <- reactive({tibble(depression = input$depression_I,
                           bmi = input$bmi_I,
                           difficult_move_objects = input$move_I,
                           snort = input$snort_I,
                           crp = input$crp_I,
                           anxious = input$anxious_I,
                           iron_um_l = input$iron_I,
                           age.cat2 = input$age_I)
  })
prob1 <- reactive({predict(log_reg_explainer, newdata = log_reg_obs())})
```


Row {data-height=450}
-----------------------------------------------------------------------
### Variable Effect on Sleeping Problem
```{r}

flexdashboard::renderValueBox({
  
flexdashboard::valueBox(value = scales::percent(round(prob1(),2)), 
         icon = "fa-bed")
})
```

### break down profile
```{r}
renderPlotly({
  plot(predict_parts_break_down(explainer = log_reg_explainer,
                                new_observation = log_reg_obs()))
})
```


Row {data-height=450}
--------------
### Vip
```{r}
vimp1 <- model_parts(log_reg_explainer)
   plot(vimp1)
```

### Performance
```{r}
```

# support vector machine

-----------------------------------------------------------------------


```{r load svm model & test data}
final_mdl_wf <- readRDS("final_svmr_wf.RDS") 
#load("ready.rda")
```


```{r xplainable, include=FALSE}
final_mdl_wf_trained <- final_mdl_wf %>% 
  fit(data = final_dat)
mdl_explainer <- explain_tidymodels(
                        model = final_mdl_wf_trained, 
                        data = final_dat %>% 
                          select(-sleep_problems), 
                        y = ready$sleep_problems,
                        label = "Radial kernel support vector machine")
```
Inputs {.sidebar}
-----------------------------------------------------------------------


Answer the following questions to find out your likelihood of having sleep problems:

```{r}
sliderInput(inputId = "bmi_j", label = "bmi",
            min = min(ready$bmi), 
            max = max(ready$bmi),
            value = 50)
sliderInput(inputId = "iron_j", label = "iron_um_l",
            min = min(ready$iron_um_l), 
            max = max(ready$iron_um_l),
            value = 30)
sliderInput(inputId = "crp_j", label = "crp",
            min = min(ready$crp), 
            max = max(ready$crp),
            value = 70)
selectInput(inputId = "snort_j", label = "snort", 
            choices = levels(ready$snort),
            selected = "never")
selectInput(inputId = "move_j", label = "difficult move objects", 
            choices = levels(ready$difficult_move_objects),
            selected = "no")
selectInput(inputId = "depression_j", label = "depression", 
            choices = levels(ready$depression),
            selected = "never")
selectInput(inputId = "anxious_j", label = "anxious", 
            choices = levels(ready$anxious),
            selected = "never")
selectInput(inputId = "age_j",  label = "age",
            choices = levels(ready$age.cat2))
```

```{r reactive expression svmr}
new_obs <- reactive({tibble(depression = input$depression_j,
                           bmi = input$bmi_j,
                           difficult_move_objects = input$move_j,
                           snort = input$snort_j,
                           crp = input$crp_j,
                           anxious = input$anxious_j,
                           iron_um_l = input$iron_j,
                           age.cat2 = input$age_j)
  })
prob <- reactive({predict(mdl_explainer, newdata = new_obs())})
```


Row {data-height=450} 
-----------------------------------------------------------------------
### Variable Effect on Sleeping Problem
```{r}
flexdashboard::renderValueBox({
  
flexdashboard::valueBox(value = scales::percent(round(prob(),2)), 
         icon = "fa-bed")
})  
```

### break down profile
```{r}
renderPlotly({
  plot(predict_parts_break_down(explainer = mdl_explainer,
                                new_observation = new_obs()))
})
```


Row {data-height=450}
--------------
### Vip
```{r}
vimp <- model_parts(mdl_explainer)
  plot(vimp)
  
```

### Performance
```{r}
```

# About
______________________________________________________________



**Background**
On average, humans spend about a 1/3 of each day sleeping. The recommended number of hours of sleep is 7 hours in a 24 hour period.^1^ Good sleep is not only vital to our day-to-day functioning but also allows us to lead healthy lives. In the US, 50-70 million people suffer from insomnia.^2^ This dashboard uses the 2018 NHANES data to explore the different variables that may have an effect on self-reported problems with sleeping.^3^ 

**Using the App**
Top variables that affect sleep were identified using Random Forest machine learning. In this app, users can explore the data and run logistic regression and different machine learning models to learn about these variables in depth.

In exploratory data analysis, users can explore the relationships between variables by changing the "X", "Y", and "Color by group" input. The logistic regression tab uses the user input data to predict the probability of reporting sleep problems given the information on age, BMI, mental and physical health, and a few laboratory values. 
The top right pane 
The top left pane
The bottom right
The bottom left

Finally, in the machine learning models, regularized logistic regression and support vector machine, the model discards all variables except for the top 3 and the predictions are made from 

**References**

[1] Watson NF, Badr MS, Belenky G, et al.; Consensus Conference Panel. Joint consensus statement of the American Academy of Sleep Medicine and Sleep Research Society on the recommended amount of sleep for a healthy adult: methodology and discussion. Sleep. 2015;38:1161–1183.

[2] Ford ES, Wheaton AG, Cunningham TJ, Giles WH, Chapman DP, Croft JB. Trends in outpatient visits for insomnia, sleep apnea, and prescriptions for sleep medications among US adults: findings from the National Ambulatory Medical Care survey 1999–2010. Sleep. 2014;37(8): 1283–1293.

[3] Centers for Disease Control and Prevention (CDC). National Center for Health Statistics (NCHS). National Health and Nutrition Examination Survey Data. Hyattsville, MD: U.S. Department of Health and Human Services, Centers for Disease Control and Prevention, [2018](https://wwwn.cdc.gov/nchs/nhanes/continuousnhanes/overview.aspx?BeginYear=2017).

[4] Packages: [Flexdashboard](https://pkgs.rstudio.com/flexdashboard/index.html), [Shiny](https://shiny.rstudio.com), 

>
>"Sleep is the swiss-army knife of health. When sleep is deficient, there is sickness and disease. And when sleep is abundant, there is vitality and health."
>
>                                                                 - Matthew Walker, PhD

**Disclaimer**
This app was created as a project for Advanced Data Science for Biomedical Engineering (EN580.464/EN580.664)/Advanced Data Science for Public Health (PH140.628/PH140.629) and is NOT validated. If you have any concerns relating to your sleep, please consult an actual healthcare provider. This app does it reflect the views of the Johns Hopkins University, the Johns Hopkins Hospital, or any of their affiliates.

**Authors**
Briha Ansari, MD., Feng-Chiao Lee, Tim Lee, MD.

**Last modified**
May 27, 2021