
load(libraries)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyverse)
library(dplyr)
library(themis)
library(RNHANES)
library(sjlabelled)
library(nhanesA)
library(ggsci)
library(ggthemes)
library(janitor)
library(rsample)
library(recipes)
library(parsnip)
library(workflows)
library(DALEX)
library(modelDown)
library(modelStudio)
```


looking for file names in each data type
```{r}

nhanesTables(data_group='DEMO', year=2018)
nhanesTables(data_group='DIET', year=2018)
nhanesTables(data_group='EXAM', year=2018)
nhanesTables(data_group='LAB', year=2018)
nhanesTables(data_group='Q', year=2018)
```


ingest data and remove SAS labels for ML models
```{r}



# demo
demo_raw <- nhanes('DEMO_J') %>% remove_all_labels()
#reading in some files of interest from the diet data
#added diet - affects sleep?
total_nutrient <- nhanes('DR1TOT_J') %>% remove_all_labels()
ind_foods <- nhanes('DR1IFF_J')  %>% remove_all_labels()
diet_supp <- nhanes('DSQTOT_J')  %>% remove_all_labels()

#reading in some files of interest from the exam data
bp <- nhanes('BPX_J')  %>% remove_all_labels()
body_measures <- nhanes('BMX_J')  %>% remove_all_labels()
dexa_femur <- nhanes('DXXFEM_J')  %>% remove_all_labels()
dexa_body <- nhanes('DXX_J')  %>% remove_all_labels()
dexa_spine <- nhanes('DXXSPN_J')  %>% remove_all_labels()
liver_US <- nhanes('LUX_J')  %>% remove_all_labels()

#reading in some files of interest from the lab data

urine_flow <- nhanes('UCFLOW_J')  %>% remove_all_labels()
fasting_q <- nhanes('FASTQX_J')  %>% remove_all_labels()
hba1c <- nhanes('GHB_J')  %>% remove_all_labels()
urine_alb_cr <- nhanes('ALB_CR_J')  %>% remove_all_labels()
uine_preg <- nhanes('UCPREG_J')  %>% remove_all_labels()
cbc <- nhanes('CBC_J')  %>% remove_all_labels()
crp <- nhanes('HSCRP_J')  %>% remove_all_labels()
total_chol <- nhanes('TCHOL_J')  %>% remove_all_labels()
hdl <- nhanes('HDL_J')  %>% remove_all_labels()
biochem <- nhanes('BIOPRO_J')  %>% remove_all_labels()
iron_status <- nhanes('FETIB_J')  %>% remove_all_labels()
LDL <- nhanes('TRIGLY_J')  %>% remove_all_labels()
cotinine <- nhanes('COT_J')  %>% remove_all_labels()
lead <- nhanes('PBCD_J')  %>% remove_all_labels()
mercury <- nhanes('IHGEM_J')  %>% remove_all_labels()
VOC <- nhanes('VOCWB_J')  %>% remove_all_labels()

##reading in some files of interest from the questionnaire data

cvs_health<- nhanes('CDQ_J')  %>% remove_all_labels()
phy_activ <- nhanes('PAQ_J')  %>% remove_all_labels()
phys_activ_youth <- nhanes('PAQY_J')  %>% remove_all_labels()
health_insurance <- nhanes('HIQ_J')  %>% remove_all_labels()
disability <- nhanes('DLQ_J')  %>% remove_all_labels()
diabetes <- nhanes('DIQ_J')  %>% remove_all_labels()
Physical_Functioning <- nhanes('PFQ_J')  %>% remove_all_labels()
recent_smoke <- nhanes('SMQRTU_J')  %>% remove_all_labels()
househol_smoke <- nhanes('SMQFAM_J')  %>% remove_all_labels()
smoking <- nhanes('SMQ_J')  %>% remove_all_labels()
med_conditions <- nhanes('MCQ_J')  %>% remove_all_labels()
current_health <- nhanes('HSQ_J')  %>% remove_all_labels()
alcohol_use <- nhanes('ALQ_J')  %>% remove_all_labels()
Aspirin_use <- nhanes('RXQASA_J')  %>% remove_all_labels()
prescrition_med <- nhanes('RXQ_RX_J')  %>% remove_all_labels()
house_char <- nhanes('HOQ_J')  %>% remove_all_labels()
derm <- nhanes('DEQ_J')  %>% remove_all_labels()
new_culture <- nhanes('ACQ_J')  %>% remove_all_labels()
sleep_disorder <- nhanes('SLQ_J')  %>% remove_all_labels()
occupattion <- nhanes('OCQ_J')  %>% remove_all_labels()
depression_screen <- nhanes('DPQ_J')  %>% remove_all_labels()
drug_use <- nhanes('DUQ_J')  %>% remove_all_labels()
childhood <- nhanes('ECQ_J')  %>% remove_all_labels()
diet_behav <- nhanes('DBQ_J')  %>% remove_all_labels()
pesticide <- nhanes('PUQMEC_J')  %>% remove_all_labels()
vol_toxicant <- nhanes('VTQ_J')  %>% remove_all_labels()
weight_history<-nhanes('WHQ_J')  %>% remove_all_labels()
mental_health <-nhanes('DPQ_J')  %>% remove_all_labels()

```

#checking for correlated data
```{r}


#alcohol use
distinct_observation <- alcohol_use %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# aspirin

distinct_observation <- Aspirin_use %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#biochem

distinct_observation <- biochem %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#body measure
distinct_observation <- body_measures %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#bp

distinct_observation <- bp %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#cbc

distinct_observation <- cbc %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#childhood

distinct_observation <- childhood %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#conitine

distinct_observation <- cotinine %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#crp
distinct_observation <- crp %>% 
distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# current health

distinct_observation <- current_health %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# cvs_health

distinct_observation <- cvs_health %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#demo

distinct_observation <- demo_raw %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#depression

distinct_observation <- depression_screen %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#derm

distinct_observation <- derm %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#dexa_body #5114

distinct_observation <- dexa_body %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#dexa femur #2898

distinct_observation <- dexa_femur %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# dexa_spine # 2898

distinct_observation <- dexa_spine %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#diabetes #8897

distinct_observation <- diabetes %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#  diet_behav #9254

distinct_observation <- diet_behav %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#diet supp #9254

distinct_observation <- diet_supp %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# diability #8897

distinct_observation <- disability %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# drug use 4572

distinct_observation <- drug_use %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation


#fasting_q #8366

distinct_observation <- fasting_q %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#hba1c #6401

distinct_observation <- hba1c %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# hdl #7435

distinct_observation <- hdl %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# health insurance #9254
distinct_observation <- health_insurance %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#house char #9254
distinct_observation <- house_char %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#household smoke #9254

distinct_observation <- househol_smoke %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#ind_foods # 112683######################
################caution: correlated data##

distinct_observation <- ind_foods %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#iron status # 6401

distinct_observation <- iron_status %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# ldl #3036

distinct_observation <- LDL %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# lead #8366

distinct_observation <- lead %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# liver # 6401

distinct_observation <- liver_US %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#med condition #8897
distinct_observation <- med_conditions %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#mental health #5533

distinct_observation <- mental_health %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# mercury # 8366

distinct_observation <- mercury %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# new culture # 8421
distinct_observation <- new_culture %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# occupation #6161

distinct_observation <- occupattion %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#pesticide # 7435

distinct_observation <- pesticide %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# phys_activ #5856

distinct_observation <- phy_activ %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation


# phys actuve youth # 2778

distinct_observation <- phys_activ_youth %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# physical functioning #8421

distinct_observation <- Physical_Functioning %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# prescription # 19643
############### caution:correlated data

distinct_observation <- prescrition_med %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# recent smoke #6401

distinct_observation <- recent_smoke %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#sleep disorder 6161

distinct_observation <- sleep_disorder %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#smoking 6724

distinct_observation <-  smoking %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# total chol 7435

distinct_observation <- total_chol %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# total nutrition # 8704

distinct_observation <- total_nutrient %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# urine preg #1057

distinct_observation <- uine_preg %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# urin alb/cr #7936

distinct_observation <- urine_alb_cr %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

#urine floe 7936

distinct_observation <- urine_flow %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# voc 3172
distinct_observation <- VOC %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# vol_toxi 3172

distinct_observation <- vol_toxicant %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation

# weight hist # 6161

distinct_observation <- weight_history %>% 
  distinct(SEQN) %>% 
  summarise(dist=n()) %>% 
  pull(dist)

distinct_observation



```
prescription meds and ind_food have correlated data (longitudinal) ie multiple records for each subject, we will not consider these 


remove redundant variables from sleep_disorder

```{r}

str(sleep_disorder)
#create sleep disorder 1 for later use
sleep_disorder1 <- sleep_disorder %>% select(c(SEQN,SLQ050,SLQ040,SLD012))

#sleep disroder with only the sqn and outcome for random forest,so that other sleep related variables dont come up during random forest feature selection 
sleep_disorder <- sleep_disorder %>% select(c(SEQN,SLQ050))
```


## combine data demo,alcohol_use,aspirin_use,biochem,body_measures,sleep_disorder, bp
```{r}

combined <- demo_raw %>%
  left_join(alcohol_use) %>% 
  left_join(Aspirin_use) %>% 
  left_join(biochem) %>% 
  left_join(body_measures) %>% 
  left_join(bp) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 10)
```


create a dataset with very imp variables

```{r}
vip1 <- combined %>% select(c(SEQN,BMXWAIST,RIDAGEYR,BMXBMI,WTINT2YR))
str(vip1)

vip1$RIDAGEYR <- as.numeric(vip1$RIDAGEYR)
str(vip1)

```


## combine data cbc,childhood,conitine,crp,current_health
```{r}

combined <- cbc %>%
  left_join(childhood) %>% 
  left_join(cotinine) %>% 
  left_join(crp) %>% 
  left_join(current_health) %>% 
  left_join(sleep_disorder)
summary(combined$LBXHSCRP)
```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 25)
```
Create vip2
```{r}
vip2 <- combined %>% select(c(SEQN,LBXHSCRP, LBXPLTSI,LBXMCVSI,LBXRBCSI,LBXRDW,LBXLYPCT,LBXNEPCT,LBXMOPCT,LBXMPSI,LBXHCT,LBXEOPCT,LBXMCHSI,HSD010,LBXWBCSI,
                              LBXMC,LBXHGB,LBDNENO,LBDLYMNO,LBXBAPCT,LBDMONO))
str(vip2)


```

## combine data cvs health,dexa_body,dexa_femur,dexa_spine,diabetes
```{r}

combined <- cvs_health %>% 
  left_join(dexa_body) %>% 
  left_join(dexa_femur) %>% 
  left_join(dexa_spine) %>% 
  left_join(diabetes) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 10)
```

## combine data diet_behav,diet_supp,disability,drug_use,fasting_q,hba1c
```{r}

combined <- diet_behav %>%
  left_join(diet_supp) %>% 
  left_join(disability) %>% 
  left_join(drug_use) %>% 
  left_join(fasting_q) %>% 
  left_join(hba1c) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 20)
```

create vip3
```{r}
vip3 <- combined %>% select(c(SEQN,PHAFSTMN,LBXGH,PHAFSTHR,DLQ140,DBD895,DLQ110,DLQ100,DLQ150,DSDCOUNT,DBD900,DBQ700,DSQTVB6,DBD905,DBD910,
                              DBQ197,DUQ220Q,DLQ130,DLQ170,DBQ235C))
str(vip3)
```


## combine data demo to hdl,health insurance,hoyuse_char,household_smoke,iron_status,lead
```{r}

combined <- hdl %>%
  left_join(health_insurance) %>% 
  left_join(house_char) %>% 
  left_join(househol_smoke) %>% 
  left_join(iron_status) %>% 
  left_join(lead) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 25)
```


get names for vip
```{r}
x<-rf_emdl$importance

print(x)

a<-as.data.frame(x)

str(a)

names(a)

print(a)

order(a)


b <-  a %>% 
      filter(MeanDecreaseGini >= 40) 

b

```

vip 4
```{r}
vip4 <- combined %>% select(c(SEQN,
LBDHDD,
LBDHDDSI,
HOD050,
LBXIRN,
LBDIRNSI,
LBXUIB,
LBDUIBSI,
LBDTIB,
LBDTIBSI,
LBDPCT,
LBXBPB,
LBDBPBSI,
LBXBCD,
LBDBCDSI,
LBXTHG,
LBDTHGSI,
LBXBSE, 
LBDBSESI, 
LBXBMN,LBDBMNSI))

str(vip4)

```

## combine data ldl,liver_us,liver_conditions,mental_health,mercury,newculture,
```{r}

combined <- LDL %>%
  left_join(liver_US) %>% 
  left_join(med_conditions) %>% 
  left_join(mental_health) %>% 
  left_join(mercury) %>% 
  left_join(new_culture) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 10)
```
create vip5

```{r}
vip5 <- combined %>% select(c(SEQN,DPQ030,WTSAF2YR))

str(vip5)
```




## combine data demo to occupation,pesticide,phys_activ,phys_activ,youth,phys function,recent_smoke
```{r}

combined <- occupattion %>%
  left_join(pesticide) %>% 
  left_join(phy_activ) %>% 
  left_join(phys_activ_youth) %>% 
  left_join(Physical_Functioning) %>% 
  left_join(recent_smoke) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 15)
```

get names for vip
```{r}
x<-rf_emdl$importance

print(x)

a<-as.data.frame(x)

str(a)

names(a)

print(a)

order(a)


b <-  a %>% 
      filter(MeanDecreaseGini >= 40) 

b

```
create vip6

```{r}
vip6 <- combined %>% select(c(SEQN,
OCQ180,
OCD270,
OCD395,
PAQ625,
PAD630,
PAQ670,
PAD675,
PAD680,
PFQ051,
PFQ061N,
PFQ061T,
PFQ063A,
PFQ063B))

str(vip6)

```



## combine data smoking,total chol,total_nutrients,urine_preg,urine_albumin,urine flow
```{r}

combined <- smoking %>%
  left_join(total_chol) %>% 
  left_join(total_nutrient) %>% 
  left_join(uine_preg) %>% 
  left_join(urine_alb_cr) %>% 
  left_join(urine_flow) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 10)
```



## combine data voc,vol_toxic,weight_histor
```{r}

combined <- VOC %>%
  left_join(vol_toxicant) %>% 
  left_join(weight_history) %>% 
  left_join(sleep_disorder)

```

# Split data
```{r}
split <- combined %>% initial_split(prop = 0.8, strata = SLQ050)
train <- training(split)
test <- testing(split)
```

## preprocessing
```{r}
rec <- recipe(x = train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_naomit(SLQ050) %>% 
  step_filter(SLQ050!=9) %>% 
  step_num2factor(SLQ050, levels = c("Y","N")) %>% 
  step_impute_median(all_predictors()) %>% 
  step_nzv(all_predictors())
  

rec
```

```{r}
train_prepped <- prep(x = rec, training = train)
train_baked <- bake(train_prepped, new_data = NULL)
```

## variable importance 

```{r fig.height=8}
rf_emdl <- randomForest::randomForest(SLQ050~., 
                                      data = train_baked %>% select(-SEQN))
randomForest::varImpPlot(rf_emdl, n.var = 10)
```


get names for vip
```{r}
x<-rf_emdl$importance

print(x)

a<-as.data.frame(x)

str(a)

names(a)

print(a)

order(a)


b <-  a %>% 
      filter(MeanDecreaseGini >= 40) 

b

```

create vip7

```{r}
vip7 <- combined %>% select(c(SEQN,WTSVOC2Y,LBXVXY,VTD271B,WHD010,WHD020,
WHD050,
WHD110,
WHD120,
WHD130,
WHD140,
WHQ150))

str(vip7)
```



MERGE VIPs

```{r}
data1 <- vip1 %>%
  left_join(vip2) %>% 
  left_join(vip3) %>% 
  left_join(vip4) %>% 
  left_join(vip5) %>% 
  left_join(vip6) %>% 
  left_join(vip7) %>% 
  left_join(sleep_disorder1) 
```
```{r}
str(data1)

```

subset to > 18 years
```{r}
data1$age <- as.numeric(data1$RIDAGEYR)

summary(data1$age)
sub <-subset (data1, age > 18)
summary(sub$age)

```
We have subset the data to >18, we will now run random forest again to choose our final predictors
## reformat
```{r}
library(dplyr)
library(tidyr)
library(recipes)
factors <- c("SEQN","SLQ040","HSD010","DLQ140","HOD050","DBQ197","DLQ130","DLQ170","DBQ235C","DLQ110","DLQ100","DLQ150","DBQ700","SLQ050","PFQ063B","DPQ030","PFQ051","PFQ061N","PFQ061T","PFQ063A")
contn <- c("PHAFSTMN","LBXPLTSI","PHAFSTHR","LBDPCT","LBDTIB","LBXUIB","LBXIRN","LBDHDD","DSDCOUNT","DBD900","DBD905","DBD910","DBD895","WHQ150","WHD120","WHD010","WHD020","WHD050","WHD110","VTD271B","OCQ180","OCD270","OCD395","PAQ625","PAD630","PAQ670","PAD675","PAD680","WHD140","WHD130","DUQ220Q")
#reformat data
sleep_tbl <- sub %>%
drop_na(SLQ050) %>%
mutate_at(factors, factor) %>%
mutate_at(contn, as.numeric)
str(sleep_tbl)
```

## split data
```{r}
library(rsample)
sleep_splits <- initial_split(data = sleep_tbl, strata = SLQ050, prop = 0.80)
sleep_train <- training(sleep_splits)
sleep_test <- testing(sleep_splits)
```

## preprocessing
```{r}
sleep_rec <- recipe(x = sleep_train, formula = SLQ050~.) %>% 
  update_role(SEQN, new_role = "ID") %>% 
  step_other(all_nominal_predictors(), threshold = 0.02) %>% 
  step_impute_mode(all_nominal_predictors()) %>% 
  step_impute_median(all_numeric_predictors()) %>% 
  step_nzv(all_predictors()) %>% 
  step_corr(all_numeric_predictors(),threshold = 0.8) %>% 
  prep()
sleep_rec
```

```{r}
sleep_train_baked <- bake(sleep_rec, sleep_train)
sleep_test_baked <- bake(sleep_rec, sleep_test)
```

## variable importance
```{r fig.height=8}
library(randomForest)
rfvi_mdl <- randomForest(SLQ050~., data = sleep_train_baked %>% select(-SEQN))
varImpPlot(rfvi_mdl)
selected_vars <- importance(rfvi_mdl) %>%  
  as_tibble( rownames = "variable") %>% 
  arrange (desc(MeanDecreaseGini)) %>% 
  slice_head(n = 20) %>% 
  pull(variable)
final_dat <- selected_vars
```

Data exploration and one at a time logistic regression to keep meaningul variables
```{r}
library(ggplot2)
# DLQ110 Are you taking anxiety pills?
# DLQ140 How often do you feel depressed?
ggplot(sleep_tbl, aes(x = SLQ050, fill= DLQ110)) +
           geom_bar()
ggplot(data = sleep_tbl) + 
  geom_bar(mapping = aes(x = DLQ140, y = ..prop.., group = SLQ050), stat = "count") + 
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~SLQ050)
logit <- glm(SLQ050 ~ DLQ140, data = sleep_tbl, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# HOD050 Number of rooms?
# Remove!
ggplot(data = sleep_tbl) + 
  geom_bar(mapping = aes(x = HOD050, y = ..prop.., group = SLQ050), stat = "count") + 
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~SLQ050)
```

```{r}
# BMXBMI - BMI
ggplot(data = sleep_tbl) + 
  geom_boxplot(mapping = aes(x = SLQ050, y = BMXBMI))
logit <- glm(data = sleep_tbl, SLQ050 ~ BMXBMI, family = "binomial")
summary(logit)
exp(coef(logit))
```
```{r}
# BMXWAIST 
# Waist circum in cm?
logit <- glm(data = sleep_tbl, SLQ050 ~ BMXWAIST, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# PFQ061T - push or pull large object
ggplot(data = sleep_tbl) + 
  geom_bar(mapping = aes(x = PFQ061T, y = ..prop.., group = SLQ050), stat = "count") + 
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~SLQ050)
logit <- glm(data = sleep_tbl, SLQ050 ~ PFQ061T, family = "binomial")
summary(logit)
exp(coef(logit))
```
```{r}
# DLQ150 
# Do you take medication for depression?
# Remove!! Surrogate for depression
sleep_tbl$SLQ050 = factor(sleep_tbl$SLQ050, levels=c(1,2), labels=c("Yes","No"))
logit <- glm(data = sleep_tbl, SLQ050 ~ DLQ150, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# Health problem?
# Discard! No healthy option
logit <- glm(data = sleep_tbl, SLQ050 ~ PFQ063A, family = "binomial")
summary(logit)
# Limit in amount of work?
logit <- glm(data = sleep_tbl, SLQ050 ~ PFQ051, family = "binomial")
summary(logit)
exp(coef(logit))
```
Outliers in age interfering with estimate, we will later categorize age in 4 bins to  get age bracket specific estimates
```{r}
# age
logit <- glm(data = sleep_tbl, SLQ050 ~ age, family = "binomial")
summary(logit)
exp(coef(logit))
plot(logit)
ggplot(data = sleep_tbl) + 
  geom_bar(mapping = aes(x = age, y = ..prop.., group = SLQ050), stat = "count") + 
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~SLQ050)
young <- sleep_tbl %>% filter(age <= 40)
old <- sleep_tbl %>% filter(age > 40)
sleep_tbl$age1 <- ifelse(sleep_tbl$age >= 40, 0, 1)
logit <- glm(data = sleep_tbl, SLQ050 ~ age1, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# SLQ040 
# How often do you snore or stop breathing?
logit <- glm(data = sleep_tbl, SLQ050 ~ SLQ040, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# DPQ030
# Trouble sleeping, sleeping too much?
logit <- glm(data = sleep_tbl, SLQ050 ~ DPQ030, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# PFQ063A
# Health problem causing difficulty
# REMOVE!
logit <- glm(data = sleep_tbl, SLQ050 ~ PFQ063A, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LBXHSCRP
# CRP?
logit <- glm(data = sleep_tbl, SLQ050 ~ LBXHSCRP, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LBDBMNSI
# Blood manganese
# Remove!! Odds 1.0017
logit <- glm(data = sleep_tbl, SLQ050 ~ LBDBMNSI, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# DLQ100
# How often do you feel worried or anxious?
logit <- glm(data = sleep_tbl, SLQ050 ~ DLQ100, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LBXPLTSI
# Plt count
# p > 0.05
logit <- glm(data = sleep_tbl, SLQ050 ~ LBXPLTSI, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LPXBPB
# Blood lead
# p > 0.05
logit <- glm(data = sleep_tbl, SLQ050 ~ LBXBPB, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LBXLYPCT
# Lymphocyte
logit <- glm(data = sleep_tbl, SLQ050 ~ LBXLYPCT, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LBXBSE
# Blood selenium
# p > 0.05
logit <- glm(data = sleep_tbl, SLQ050 ~ LBXBSE, family = "binomial")
summary(logit)
exp(coef(logit))
```

```{r}
# LBDIRNSI
# Blood iron
logit <- glm(data = sleep_tbl, SLQ050 ~ LBDIRNSI, family = "binomial")
summary(logit)
exp(coef(logit))
```
Below are the finalists
```{r}
final_dat <- sleep_tbl %>% select(c(SLQ050,
                                    DLQ140, 
                                    BMXWAIST, 
                                    BMXBMI, 
                                    PFQ061T,
                                    DLQ150,
                                    age,
                                    SLQ040,
                                    LBXHSCRP,
                                    DLQ100,
                                    LBDIRNSI))
```


```{r}
library(recipes)
library(lares) # for correlation check
library(forcats)
library(tidyverse)
dat2 <- final_dat


summary(dat2)
```

# remiove 7 and 9

```{r}
dat2 <- dat2 %>% 
        filter(SLQ050 != 9) %>% 
        filter(DLQ140 != 9) %>% 
        filter(DLQ140 != 7) %>% 
        filter(PFQ061T != 9) %>% 
        filter(PFQ061T != 7) %>% 
        filter(DLQ150 != 7) %>% 
        filter(DLQ150 != 9) %>% 
        filter(SLQ040 != 7) %>% 
        filter(SLQ040 != 9) %>% 
        filter(DLQ100 != 7) %>% 
        filter(DLQ100 != 9) 

table(dat2$SLQ040)

summary(dat2)

str(dat2)

summary(dat2$LBXHSCRP)
summary(dat2$LBDIRNSI)
```

```{r}
#reformat to as.factor and as.numeric
factors <- c("SLQ050","DLQ140","DLQ100","DLQ150","SLQ050","PFQ061T","SLQ040")


#reformat data
dat2 <- dat2 %>%
mutate_at(factors, factor) 
       
dat2$age <- as.numeric(dat2$age)
str(dat2)

hist(dat2$age)

                    
summary(dat2)

```

missing data

```{r}

#check missing data
sapply(dat2, function(x) sum(is.na(x)))

library(mice)
init = mice(dat2, maxit=0) 
meth = init$method
predM = init$predictorMatrix
```

```{r}
library(VIM)
aggr_plot <- aggr(dat2, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

To impute the missing values, mice package use an algorithm in a such a way that use information from other variables in the dataset to predict and impute the missing values. Therefore, you may not want to use a certain variable as predictors. For example, the ID variable does not have any predictive value.

```{r}
view(dat2)
```

no id variable so nothing to remover, we will use predictive mean matching to impute since "norm" gave -ive values.Predictive mean matching calculates the predicted value of target variable Y according to the specified imputation model. For each missing entry, the method forms a small set of candidate donors (typically with 3, 5 or 10 members) from all complete cases that have predicted values closest to the predicted value for the missing entry. One donor is randomly drawn from the candidates, and the observed value of the donor is taken to replace the missing value. The assumption is the distribution of the missing cell is the same as the observed data of the candidate donors, source : https://stefvanbuuren.name/fimd/sec-pmm.html
```{r}
meth[c("BMXWAIST")]="pmm" 
meth[c("BMXBMI")]="pmm" 
meth[c("LBXHSCRP")]="pmm" 
meth[c("LBDIRNSI")]="pmm"

set.seed(103)
imputed = mice(dat2, method=meth, predictorMatrix=predM, m=5)

imputed <- complete(imputed)

sapply(imputed, function(x) sum(is.na(x)))

summary(imputed)

methods(mice)
```
recheck missingness
```{r}
aggr_plot <- aggr(imputed, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(data), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

making sure imputed data has no negative numbers
```{r}

summary(imputed$LBXHSCRP)
summary(imputed$LBDIRNSI)
summary(imputed$BMXBMI)
summary(imputed$BMXWAIST)
```
check for correlated data, and drop meaningless variables
```{r}
corr_cross(imputed, # name of dataset
  max_pvalue = 0.05, # display only significant correlations (at 5% level)
  top = 10 # display top 10 couples of variables (by correlation coefficient)
)

corr_cross(imputed)
```
we will drop DLQ150, BMXWAIST
```{r}
clean_baby <- imputed %>% select(-DLQ150, -BMXWAIST)
```


```{r}
dat <- clean_baby

names(dat)

levels(clean_baby$SLQ050)
table(clean_baby$SLQ050)
```



recode SLQ050 to keep no as a reference and rename to sleep_problems, rename all variables
```{r}
dat1 <- dat %>%
mutate(SLQ050 = relevel(SLQ050, ref = "No")) %>%
rename(sleep_problems = SLQ050)  %>%
mutate(DLQ140 = fct_recode(.f = DLQ140, "Daily" = "1", "weekly" = "2", "monthly" = "3", "few_times_year" = "4", "never" = "5")) %>%
rename(depression = DLQ140) %>%
rename(bmi = BMXBMI) %>%
mutate(PFQ061T = fct_recode(.f = PFQ061T, "no" = "1", "some" = "2", "much" = "3", "unable" = "4", "dont_do_this_activity" = "5")) %>%
rename(difficult_move_objects = PFQ061T ) %>%
mutate(SLQ040 = fct_recode(.f = SLQ040, "never" = "0", "1-2/wk" = "1", "3-4/wk" = "2", "5/wk" = "3")) %>%
rename(snort = SLQ040) %>%
rename(crp = LBXHSCRP) %>%
mutate(DLQ100 = fct_recode(.f = DLQ100, "daily" = "1", "weekly" = "2", "monthly" = "3", "few/yr" = "4", "never" = "5")) %>%
rename(anxious = DLQ100) %>%
rename(iron_um_l = LBDIRNSI)


table(dat1$sleep_problems)
```

check variable freq after recoding
```{r}
# sleep_problems
table(dat$SLQ050)
table(dat1$sleep_problems)
# depression
table(dat$DLQ140)
table(dat1$depression)
# BMI
summary(dat$BMXBMI)
summary(dat1$bmi)
# difficulty moving large objects
table(dat$PFQ061T)
table(dat1$difficult_move_objects)
# snort
table(dat$SLQ040)
table(dat1$snort)
# crp
summary(dat$LBXHSCRP)
summary(dat1$crp)
# anxious
table(dat$DLQ100)
table(dat1$anxious)
# crp
summary(dat$LBDIRNSI)
summary(dat1$iron_um_l)


str(dat1)
```
categorizing age into 4 bins
```{r}
dat1$age.cat2 <- cut(dat1$age, breaks = 4, labels = c("19-35", "36-50", "51-65","above65"))


# 19-36 level1, 36-51 level2, 51- 65 level 3, >= 66 level 4

table(dat1$age.cat2)
```
remove age and save file as ready for models and app
```{r}


ready <- dat1 %>% select(-age)
save(ready, file="ready.rda")

str(ready)
```



